---
title: "MFO renewal sorting"
author: "Peter Smalldige"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

test
git checkout --orphan clean-main
git add .

```{r read data, echo=F, include=T, message=F, eval=T}
# library(ggplot2)
# library(ggpubr)
# library(expss)
library(readr)
library(readxl)
library(writexl)
# library(skimr)
 library(tidyverse)  # includes: tidyr, dplyr, haven, readr, readxl, forcats, stringer, ggplot2. See all with tidyverse_packages()
# library(knitr)
# library(janitor)

```


```{r echo = FALSE}
# Create Paths
file_path_input <- "C:\\Users\\pjs23\\Documents\\R\\MFO_Annual_Summary\\Renewal data\\renewal_2025\\database_2024.xlsx"
#initil commit to git on 15 april 2025

output_folder <- "C:\\Users\\pjs23\\Documents\\R\\MFO_Annual_Summary\\Renewal data\\renewal_2025"

```

Baseline, last year's volunteer list
```{r read data, echo=F, include=T, message=F, eval=T}

#the data from the 2024 database of active and inactive volunteers. The list of last year's volunteers.
vol_last_year <- read_excel(file_path_input, sheet = "volunteers_2024", skip=0) %>%
  mutate(
    count =1,
    email = tolower(email),
    source = "original"
  ) %>% 
  mutate(email = str_remove_all(email, "[; ]")) %>% #removes blanks and ; from the email varaible
  select(first, last, email, source, count, region) %>% 
  arrange(email, region) %>% 
  group_by(email) %>% 
  summarize(
    first = first(first),
    last = first(last),
    email = first(email),
    source = first(source),
    count = first(count),
    region = first(region)
  ) %>% 
  mutate(
    region = str_to_lower(region),  # Convert to lowercase
    region = case_when(
      str_detect(region, "capital") ~ "capital/E_ADK",
      str_detect(region, "southern") ~ "southern",
      str_detect(region, "northwest") ~ "northwest",
      str_detect(region, "central") ~ "central/W_ADK",
      str_detect(region, "catskill") ~ "catskill/L_HDS",
      TRUE ~ region  # Keep as is if no match
    )
  )


```

```{r read data, echo=F, include=T, message=F, eval=T}

# Obtain the responses from the qualtrics survey. Export and copy into a new sheet named "renewed_2025". 
# Update the worksheet from the Qualtrics output file. Save the previous worksheet as a backup, or delete.
renewed <- read_excel(file_path_input, sheet = "renewed_2025", col_names = FALSE, skip = 4) %>% 
 rename(
    date = 1,
    first = 2,
    last = 3,
    status = 4,
    apprehension = 5,
    county = 6,
    no_additional = 7,
    #additional county list columns 8 - 69
    region = 70,
    address = 71,
    property_address = 72,
    phone_home = 73,
    phone_cell = 74,
    landline = 75,
    cell_line = 76,
    email1 = 77,
    email2 = 78,
    email_spouse = 79
  ) %>%
  mutate(email1 = str_remove_all(email1, "[; ]")) %>% #removes blanks from the email1 variable
  mutate(email2 = str_remove_all(email2, "[; ]")) %>% #removes blanks from the email2 variable
  mutate(email_spouse = str_remove_all(email_spouse, "[; ]")) %>% #removes blanks from the email_spouse variable
  mutate(
    issue = "",
    count = 1,
    source = "responded",
    email1 = tolower(email1),
    email2 = tolower(email2),
    email_spouse = tolower(email_spouse)
    ) %>% 
  #filter(! is.na(email1)) %>%  
  mutate(
    issue = case_when(
      email1 == email2 ~ "none",
      TRUE ~ "yes"
  )) %>% 
  rename(email = email1) %>% 
  mutate(
    region = str_to_lower(region),  # Convert to lowercase
    region = case_when(
      str_detect(region, "capital") ~ "capital/E_ADK",
      str_detect(region, "southern") ~ "southern",
      str_detect(region, "northwest") ~ "northwest",
      str_detect(region, "central") ~ "central/W_ADK",
      str_detect(region, "catskill") ~ "catskill/L_HDS",
      TRUE ~ region  # Keep as is if no match
    )
  )


renewed2 <- renewed %>% # a simple version for comparison to the 2024 list. Use the long df for use in 2025. Will need to make it tidy.
  select(first, last, email, email_spouse, source, count, region)

bad_email <- renewed %>% 
  filter(issue == "yes")

```


```{r}

# name the clusters of columns for future reference

 countynames <- names(renewed)[8:69]
 phonedetails <- names(renewed)[104:108]
# distance <- names(renewed)[109]
 mgmtlist <- names(renewed)[110:134]
 textdetails <- names(renewed)[99:103]
 emaildetails <- names(renewed)[94:98]
 contacting <- names(renewed)[88:93]
 acre_range <- names(renewed)[81:86]
 
# change the unlabeled columns to something meaningful so the regional directors
# can easily access the data for the current volunteers.
# 
 
final <- renewed %>% 
  #albany = col 08, yates = col 69, "no additional counties" = col 07

  rename(
    numb_visits = '...80',
    team_pref = '...87',
    distance_pref = '...109',
    other_info = "...135",
    land_line = landline
  ) %>% 

  
 mutate(across(countynames, as.character)) %>%        # Convert columns 8–69 to character
 unite("addl_counties", all_of(countynames), sep = ", ", na.rm = TRUE) %>%   # Concatenate, skipping NAs. Unite drops the original counties

 # mutate(across(acre_range, as.character)) %>% 
 #  unite("acre_pref", all_of(acre_range), sep = "; ", na.rm = TRUE) %>% 
  
# separate out the smallest and largest of acre preferences, and ignore those in the middle.
# Concatenate with "x to y"


  mutate(across(all_of(acre_range), ~ str_trim(as.character(.)))) %>%  # trim whitespace
  unite("acre_pref", all_of(acre_range), sep = "; ", na.rm = TRUE) %>%

  
mutate(
  acre_pref_split = str_split(acre_pref, "; "),
  acre_pref_range = map_chr(acre_pref_split, ~ {
    values <- na.omit(.x) %>% str_trim()
    if (length(values) == 0) return(NA_character_)
    if (length(values) == 1) return(values)
    paste0("(", values[1], ") to (", values[length(values)], ")")
  })
) %>%     
    
    # acre_pref_range = map_chr(acre_pref_split, ~ {
    #   values <- na.omit(.x) %>% str_trim()
    #   if (length(values) == 0) return(NA_character_)
    #   if (length(values) == 1) return(values)
    #   paste0(values[1], " to ", values[length(values)])
    # })

  
 # mutate(across(distance, as.character)) %>% 
 #  unite("distance_trvl", all_of(distance), sep = "; ", na.rm = TRUE) %>% 

 mutate(across(phonedetails, as.character)) %>% 
  unite("phone_check", all_of(phonedetails), sep = "; ", na.rm = TRUE) %>% 
  
 mutate(across(textdetails, as.character)) %>% 
  unite("text_check", all_of(textdetails), sep = "; ", na.rm = TRUE) %>% 
  
 mutate(across(emaildetails, as.character)) %>% 
  unite("email_check", all_of(emaildetails), sep = "; ", na.rm = TRUE) %>% 
  
 mutate(across(contacting, as.character)) %>% 
  unite("contact_me", all_of(contacting), sep = "; ", na.rm = TRUE) %>% 
  
  mutate(across(mgmtlist, as.character)) %>% 
  unite("mgmt_expr", all_of(mgmtlist), sep = "; ", na.rm = TRUE) %>% 
  

    # this code isn't needed.  The "unite" function removes the columns used to merge
  # select(additional_counties, everything()) %>% 
  #   select(-c('...110': '...134')) %>% 
  #   select(-c('...104': '...108')) %>% 
  #   select(-c('...99':'...103')) %>% 
  #   select(-c('...94': '...98')) %>% 
  #   select(-c('...88': '...93')) %>% 
  #   select(-c('...81': '...86')) %>% 
  #   select(-c(`...8`:`...69`)) %>% 

  
  mutate(
    land_line = str_remove_all(land_line, "\\D"),  # Remove all non-numeric characters
    land_line = str_replace(land_line, "(\\d{3})(\\d{3})(\\d{4})", "(\\1) \\2-\\3")  # Format as (XXX) XXX-XXXX
  ) %>% 
  
  mutate(
    cell_line = str_remove_all(cell_line, "\\D"),  # Remove all non-numeric characters
    cell_line = str_replace(cell_line, "(\\d{3})(\\d{3})(\\d{4})", "(\\1) \\2-\\3")  # Format as (XXX) XXX-XXXX
  ) %>% 
  
    
  select(date, first, last, region, status, apprehension, county, addl_counties,
         address, property_address, contact_me, land_line, cell_line, email, email_spouse, 
         phone_check, text_check, email_check, numb_visits, 
         team_pref, distance_pref, acre_pref, acre_pref_range, 
         mgmt_expr, other_info) %>% 
  arrange(status, region, county, last)
  

```




```{r}
#change the name of the output file to today's date
output_file <- file.path(output_folder, paste0("MFO_data_14april2025", ".xlsx"))

write_xlsx(final, output_file)
```





1. Identify respondents
2. Remove respondents from the original request.
3. Write file of non_respondents and resend email.

```{r read data, echo=F, include=T, message=F, eval=T}

#sort and filter respondents based on primary email address
respondents <- bind_rows(renewed2, vol_last_year) %>% 
  arrange(email, desc(source)) %>% # sorts SOURCE=original before SOURCE=responded, so reverse sort "source"
  group_by(email) %>% 
  summarize(
    region = first(region),
    source = first(source), #keeps first value of source=original
    first = first(first), # keeps "first" name
    last = first(last), #keeps "last" name 
    email_spouse = first(email_spouse),
    total = sum(count), #for those in the emailed and response data, total will equal 2
    .groups = "drop"
    )

#sort and filter respondents based on email_spouse address
spouse <- bind_rows(renewed2, vol_last_year) %>%
  arrange(email, desc(source)) %>% # sorts SOURCE=original before =responded
  group_by(email) %>% 
  summarize(
    region = first(region),
    source = first(source), #keeps first value of source=original
    first = first(first), # keeps "first" name
    last = first(last), #keeps "last" name 
    total = sum(count), #for those in the emailed and response data, total will equal 2
    email_spouse = first(email_spouse),
    .groups = "drop"
    ) %>% 
  #remove when total =2, drop if email_spouse is NA
  filter(!(total == 2 & is.na(email_spouse))) %>% 
  filter(!(email == email_spouse)) %>% # removes rows where the volunteer included their email as "email for spouse"
  select(!c(email, total, source)) %>%
  select(first, last, email_spouse) %>% 
  rename(email = email_spouse) %>% 
  mutate(
    source = "spouse",
    count = 1
  ) %>% 
  bind_rows(vol_last_year) %>%     # add back the original volunteer list
  #   arrange(email, desc(source)) %>%     # sorts SOURCE=spouse before =original
  # group_by(email) %>% 
  filter(source == "spouse")

lastname <- bind_rows(renewed2, vol_last_year) %>%
  arrange(last, desc(source)) %>% # sorts SOURCE=original before =responded
  group_by(last) %>% 
  summarize(
    region = first(region),
    email = first(email),
    source = first(source), #keeps first value of source=original
    first = first(first), # keeps "first" name
    last = first(last), #keeps "last" name 
    total = sum(count), #for those in the emailed and response data, total will equal 2
    .groups = "drop"
    ) %>% 
  filter(total > 1) %>% 
  select(email, source, first, last, region)


 # merge df=respondents with df=spouse. df=respondents includes responded and original. "original" not yet responded.
 # Removing those who responded retains spouse and original. Combine spouse and original, filter those where total >1
 # some volunteers listed a secondary email in the "spouse" field (e.g., Goff, Farber). Recode those to responded.
 # total the count for group_by email and remove total > 1 and "responded"  Residual data are non respondents.
 # 
no_response <- bind_rows(respondents, spouse) %>% 
  select(!c(email_spouse, total, count)) %>% 
 # filter(!(source == "responded" )) %>% 
  mutate(source = if_else(source == "spouse", "responded", source)) %>%  #if source = "spouse" then change source to "responded"
      #this is for people who have two emails, one they added to the spouse field e.g., Goff, Farber. Filter below removes those who  "responded"
  ungroup() %>% 
  mutate(
    count = 1, 
    email = tolower(email)
  ) %>% 
  arrange(email, desc(source)) %>% 
  group_by(email) %>% 
  summarize(
    region = first(region),
    source = first(source),
    first = first(first),
    last = first(last),
    total = sum(count),
    .groups = "drop"
  ) %>% 
  filter(!(total > 1 | source == "responded")) %>% 
#  filter(!(email == "janisby@gmail.com" | email == "troy.weldy@clctrust.org")) %>% 

# bind rows of lastname to extract respondents who changed their email address between last year and this year (e.g., Walt Williamson)
  bind_rows(lastname) %>% 
  mutate(
    count = 1,
    email = tolower(email)
  ) %>% 
  select(-total) %>% 
  arrange(last, desc(source)) %>% 
  group_by(last) %>% 
  summarize(
    region = first(region),
    source = first(source),
    first = first(first),
    last = first(last),
    email = first(email),
    total = sum(count),
    .groups = "drop"
  ) %>% 
  #Remove if source = "responded". The use of "total" will capture parent/child volunteers (e.g., Aujero) who have not responded.
  #From the df=lastname, the source == "responded" for those who changed their email address (e.g., Williamson)
  filter(!(source=="responded")) %>% 
  select(region, email, first, last) %>% 
  arrange(region,last)


```

```{r}
#change the name of the output file to today's date
output_file <- file.path(output_folder, paste0("nonrespondents11april2025", ".xlsx"))

write_xlsx(no_response, output_file)
           
```



# Construct the full file path
output_file_ <- file.path(output_folder, paste0(webinar_name, "_CEU_1.xlsx"))

# Write the CEU_1 data frame to the specified path
write_xlsx(CEU_1, output_file_ceu)


# Construct the full file path
output_file_ceu <- file.path(output_folder, paste0(webinar_name, "_CEU_1.xlsx"))

# Write the CEU_1 data frame to the specified path
write_xlsx(CEU_1, output_file_ceu)


Final Sorting...determine who needs to be called.

```{r read data, echo=F, include=T, message=F, eval=F}

# combine the nonrespondent data with the original data file to 
# indicate who needs to be called

nonrespondents2 <- nonrespondents %>% 
  select(email, source, count)


missing <- bind_rows(nonrespondents2, active) %>% 
  arrange (source) %>% 
  group_by(email ) %>% 
  summarize(
    #source = first(source), #keeps first value of source=active and replaces the alternate value of
                             # subsequent levels of the value of source
    total = sum(count), #for those in the emailed and response data, total will equal 2
    across (everything())
  ) %>% 
  filter (total >1) %>% 
  arrange(desc(source)) %>% 
  summarize(
    source= first(source),
    across (everything())
  ) %>% 
  drop_na(first) %>% 
  select(region, source, county, first, last, homephone, cellphone, email)

write.csv(missing, "missing-actives_6march.csv")

```

